---
title: 'West Vancouver Library - Preliminary Outdoor and Indoor AQ Data'
date: "`r Sys.Date()`"
output:
  pdf_document:
    number_sections: no
    toc: yes
    fig_caption: yes
  word_document:
    toc: no
  html_document:
    df_print: paged
    toc: yes
authors: Hugo Dignoes Ricart, Caroline Webber, Krista Traboulay, and Prof. Naomi Zimmerman
---

\newpage

```{r setup, include=FALSE}
library(patchwork)
library(rmarkdown)
library(openair)
library(gtools)
library(ggplot2)
library(caret)
library(plyr)
library(dplyr)
library(readr)
library(purrr)
library(tidyr)
library(kableExtra)
library(tidyverse)
library(captioner)
library(lubridate)
library(lattice)
library(latticeExtra)
library(worldmet)
library(readxl)
library(grid) # For legend grob
library(gridExtra) # For extra legend when plotting
library(png) # For forcing legend to centre
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
set.seed(1703)
fig_cap <- captioner::captioner()
tab_cap <- captioner::captioner("Table")
figure_nums <- captioner(prefix = "Figure")
```

```{r Adjustable Parameters, include = FALSE, message = FALSE, warning = FALSE}
month_char <- "August"
month_int <- match(month_char, month.name) # Keep as is
year <- 2025
```

```{r Load and Subset Data, include = FALSE, message = FALSE, warning = FALSE}
# Load data
outdoor_data <- read.csv("other_data/2021_outdoor_Aug_data_with_AQHI.csv")
indoor_data <- read.csv("other_data/2020_indoor_Aug_data_with_AQHI.csv")

data_list <- list(outdoor_data, indoor_data)

for (i in seq_along(data_list)) {
  # Remove rows with NAs
  data_list[[i]] <- na.omit(data_list[[i]])

  # Ensure consistent date formatting (not needed if dates in UTC)
  data_list[[i]]$date <- ifelse(
    grepl(":", data_list[[i]]$date),
    data_list[[i]]$date,
    paste0(data_list[[i]]$date, " 00:00") # add midnight if missing
  )

  data_list[[i]]$date <- as.POSIXct(data_list[[i]]$date, format = "%m/%d/%Y %H:%M", tz = "Etc/GMT-8")

  # Rename columns
  names(data_list[[i]])[names(data_list[[i]]) == "NO2..ppb."] <- "NO2"
  names(data_list[[i]])[names(data_list[[i]]) == "NO..ppb."] <- "NO"
  names(data_list[[i]])[names(data_list[[i]]) == "CO2..ppm."] <- "CO2"
  names(data_list[[i]])[names(data_list[[i]]) == "O3..ppb."] <- "O3"
  names(data_list[[i]])[names(data_list[[i]]) == "CO..ppm."] <- "CO"
  names(data_list[[i]])[names(data_list[[i]]) == "PM2_5.ug.m3."] <- "PM2_5"
}

outdoor_data <- data_list[[1]]
indoor_data <- data_list[[2]]

# Define pollutants and their units
pollutants <- c("NO2", "NO", "CO2", "O3", "CO", "PM2_5")

units <- list(
  NO2 = "ppb",
  NO = "ppb",
  CO2 = "ppm",
  O3 = "ppb",
  CO = "ppm",
  PM2_5 = "µg/m³"
)

# Hourly averages
outdoor_data_hr_avg <- timeAverage(outdoor_data, avg.time = "hour")
indoor_data_hr_avg <- timeAverage(indoor_data, avg.time = "hour")

# 24 hr averages
outdoor_data_24hr_avg <- timeAverage(outdoor_data, avg.time = "day")
indoor_data_24hr_avg <- timeAverage(indoor_data, avg.time = "day")

# Round AQHI after taking average
outdoor_data_24hr_avg$AQHI <- round(outdoor_data_24hr_avg$AQHI)
indoor_data_24hr_avg$AQHI <- round(indoor_data_24hr_avg$AQHI)
```

```{r Load plotting functions, include = FALSE, message = FALSE, warning = FALSE}
# Formatted variables are expressions with subscripts/ special chars
generate_timeseries_plot <- function(
    pollutant_name, y_label_expression,
    outdoor_data_hr_avg, indoor_data_hr_avg) {
  # Get ranges first
  xrange <- range(c(outdoor_data_hr_avg$date, indoor_data_hr_avg$date))
  yrange <- range(
    c(
      outdoor_data_hr_avg[[pollutant_name]],
      indoor_data_hr_avg[[pollutant_name]]
    )
  )
  # Empty plot
  plot(NA,
    xlim = xrange, ylim = yrange,
    xlab = "Date",
    ylab = y_label_expression,
    xaxt = "n"
  )
  # Add title
  title(main = bquote(
    "Outdoor and Indoor" ~ .(y_label_expression[[1]]) ~ "vs Time"
  ))
  # Add grid lines
  grid(col = "#b9b9b9", lty = 1)
  # Add lines
  lines(
    outdoor_data_hr_avg$date, outdoor_data_hr_avg[[pollutant_name]],
    col = "blue"
  )
  lines(indoor_data_hr_avg$date,
    indoor_data_hr_avg[[pollutant_name]],
    col = "red"
  )
  # Format axis
  axis.POSIXct(
    1,
    indoor_data_hr_avg$date, # Assumes same dates in outdoor and indoor data
    format = "%b %d"
  )
  # Add legend
  legend(
    x = "topright", box.col = "black",
    bg = "#fafaa4", box.lwd = 2,
    legend = c(
      "Outdoor Concentrations",
      "Indoor Concentrations"
    ),
    fill = c("blue", "red")
  )
}


generate_pollutant_calendar_plot <- function(
  dataset, dataset_col_name, year_int, month_int,
  title_expression, threshold_value, colours
) {
  calendarPlot(
    dataset, pollutant=dataset_col_name, year=year_int,
    month=month_int, limits = c(0, threshold_value),
    cols = colours,
    main = title_expression
  )
}

generate_AQHI_calendar_plot <- function(
  dataset, dataset_col_name, year_int, month_int,
  is_outdoor_data
) {
  if (is_outdoor_data) {
    title <- "Outdoor Air Quality Health Index"
  } else {
    title <- "Indoor Air Quality Health Index"
  }
  key_labels <- c("Low - 1", "Low - 2", "Low - 3", "Moderate - 4", "Moderate - 5",
    "Moderate - 6", "High - 7", "High - 8", "High - 9",
    "High - 10", "Extreme - 10+"
  )

  colours <- c("#00ccff", "#0099cc", "#006699", "#ffff00", "#ffcc00",
    "#ff9933", "#fe6666", "#ff0000", "#cc0000", "#990000", "#660000"
  )

  index_breaks <- c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,Inf)
  
  calendarPlot(
    dataset, pollutant=dataset_col_name, year=year_int,
    month=month_int, main = title, breaks = index_breaks,
    cols = colours, labels = key_labels
  )
}

# For both indoor and outdoor datasets
get_diurnal_plots <- function(
    outdoor_dataset, indoor_dataset, pollutant_name,
    y_label_expression, horizontal_references) {
  diurnal_outdoor <- timeVariation(outdoor_dataset,
    pollutant = pollutant_name,
    ylab = y_label_expression,
    xlab = "Hour", cols = "blue",
    ref.y = horizontal_references
  )[["plot"]][["hour"]]

  diurnal_indoor <- timeVariation(indoor_dataset,
    pollutant = pollutant_name,
    ylab = y_label_expression,
    xlab = "Hour", cols = "red",
    ref.y = horizontal_references
  )[["plot"]][["hour"]]

  diurnal_outdoor[["main"]] <- bquote(
    "Hourly average of outdoor" ~ .(y_label_expression[[1]])
  )
  diurnal_indoor[["main"]] <- bquote(
    "Hourly average of indoor" ~ .(y_label_expression[[1]])
  )
  return(list("outdoor_plot" = diurnal_outdoor, "indoor_plot" = diurnal_indoor))
}


# For both indoor and outdoor datasets
get_box_plots <- function(
    outdoor_dataset, indoor_dataset, pollutant_name,
    y_label_expression) {
  box_outdoor <- ggplot(
    outdoor_dataset, aes(x = "", y = !!sym(pollutant_name))
  ) +
    geom_boxplot(fill = "blue") +
    labs(
      y = y_label_expression, x = "",
      main = bquote("Outdoor" ~ .(y_label_expression[[1]]) ~ "Distribution")
    ) +
    theme_minimal()
  box_indoor <- ggplot(indoor_dataset, aes(x = "", y = !!sym(pollutant_name))) +
    geom_boxplot(fill = "red") +
    labs(
      y = y_label_expression, x = "",
      main = bquote("Outdoor" ~ .(y_label_expression[[1]]) ~ "Distribution")
    ) +
    theme_minimal()

  return(list("outdoor_plot" = box_outdoor, "indoor_plot" = box_indoor))
}

make_CAAQS_legend <- function(legend_title,
  grn_val, ylw_val1, ylw_val2, org_val1,
  org_val2, red_val) {
  op <- par(no.readonly = TRUE)
  plot(
    0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
    xlim = c(0, 1), ylim = c(0, 1)
  )
  legend(
    "center",
    title = legend_title,
    legend = c(sprintf("Green (<=%s)", grn_val), sprintf("Yellow (%s-%s)",
      ylw_val1, ylw_val2), sprintf("Orange (%s-%s)",
      org_val1, org_val2), sprintf("Red (>%s)", red_val)),
    pch = 22,
    pt.bg = c("#187a1840", "#ffea0060", "#ff800040", "#ff000040"),
    col = "black",
    pt.cex = 2,
    bty = "n", cex = 0.8, xpd = NA
  )
  par(op)
}
```

# Background and Introduction
This report provides the results of air monitoring that was conducted at your facility during August 2025.

The patterns and levels of NO, NO~2~, PM~2.5~, CO, CO~2~, and O~3~ outside and inside the library will be summarized with a series of functions and plots found in the R package ‘openair’ [D. C. Carslaw and Ropkins (2012)].

The purpose of this report is to provide the reader with a better understanding of the following:

* Reading and interpreting Hourly Average, and Calendar plots

If you have questions about this report, please email Prof. Naomi Zimmerman, Caroline Webber or Hugo Dignoes Ricart at  [HelloAirQuality@mech.ubc.ca](mailto:HelloAirQuality@mech.ubc.ca) 

# Study Instrumentation

To collect the air pollutant data at your facility, low-cost air pollutant sensors, the SENSIT RAMP (remote air quality monitoring platform) and/or the QuantAQ MODULAIR were deployed. Low-cost sensors are powerful tools that can be used to make air pollution data more accessible. However, because these sensors are lower-cost, it is possible for the sensors to fail, and calibrations must be performed on the raw data to ensure accuracy. Once the sensors are properly calibrated, their data are generally reliable. The reported data has been calibrated.

# How we are summarizing air pollution data

## Calendar Plots

A calendar plot is a visualization of the 24-hour average concentrations of a pollutant over the course of the deployment. The dates on the calendar are coloured according to the relative concentrations of the pollutant. The dates are coloured on a gradient, with the darker reds corresponding to measured concentrations approaching the exposure limits set by regulatory bodies.

## Hourly Average Plots

When analyzing air pollution data, useful information can be deducted from the typical pattern of air pollution variations throughout the course of a day. An hourly average plot graphs the average concentration of a monitored pollutant at each hour of the day. Variations in the pollutant’s concentration can be used to interpret its source(s). It is important to note that each point on the daily average plot represents an average across all days for the specified hour. For example, the point at 6AM would represent the average concentration between 5-6AM over the complete period of monitoring. Day-to-day concentration variations are represented by shaded regions above or below the data point, where more day-to-day variability will be represented with larger shaded regions.

## Units of Measurement
Air pollutant concentrations are reported here with three units: 

1. **Part per billion (ppb)**: This is the volume fraction of the air pollutant in the air. Think of this the same way you would think of a "percent". One percent as a fraction is 1/100 = 0.01. One part per billion (ppb) is 1/1,000,000,000 = $1\times10^{-9}$. So if a pollutant concentration is 25 ppb, then that means there is $25\times10^{-9}$ m$^{3}$ of the pollutant in one cubic metre of air. We use this unit to report concentrations of nitrogen oxides (NO~x~), and ozone (O~3~).

2. **Part per million (ppm)**: This is the volume fraction of the air pollutant in the air. Think of this the same way you would think of a "percent". One percent as a fraction is 1/100 = 0.01. One part per million (ppm) is 1/1,000,000 = $1\times10^{-6}$. So if a pollutant concentration is 25 ppm, then that means there is $25\times10^{-6}$ m$^{3}$ of the pollutant in one cubic metre of air. We use this unit to report concentrations of carbon monoxide (CO) and carbon dioxide (CO~2~).

3. **Micrograms per cubic metre (µg/m$^{3}$)**: This is a mass fraction and is used for measuring the mass of particulate matter in the air. One microgram is equivalent to $1\times10^{-6}$ grams. So if a pollutant concentration is 25 µg/m$^{3}$, then that means there is $25\times10^{-6}$ g of the pollutant in one cubic metre of air. We use this unit to report concentrations of PM~2.5~ (Particles with a diameter less than 2.5 microns).



\newpage

# Air Quality Health Index (AQHI)

## What is the Air Quality Health Index (AQHI)? *[paraphrased from Health Canada]*

The Air Quality Health Index or “AQHI” is a scale designed to help you understand what the air quality
around you means to your health.
It is a health protection tool that is designed to help you make decisions to protect your health by limiting
short-term exposure to air pollution and adjusting your activity levels during increased levels of air pollution.
It also provides advice on how you can improve the quality of the air you breathe.
This index pays particular attention to people who are sensitive to air pollution and provides them with
advice on how to protect their health during air quality levels associated with low, moderate, high and very
high health risks.

## How is the AQHI calculated?

In British Columbia, the AQHI is calculated concurrently in two ways, and the highest of the two values is
reported:

1. A calculation based on the relative risks of a combination of common air pollutants that is known to
harm human health:
    a. Ozone (O~3~) at ground level,
    b. Particulate Matter (PM~2.5~) and
    c. Nitrogen Dioxide (NO~2~).
2. Particulate Matter (PM~2.5~) concentration alone

## What is the scale for the AQHI?

The AQHI is measured on a scale ranging from 1-10+. The AQHI index values are grouped into health risk
categories as shown below. These categories help you to easily and quickly identify your level of risk.

* 1-3 Low health
* 4-6 Moderate health risk
* 7-10 High health risk
* 10 + Very high health risk

From the data collected at your facility, we calculated the AQHI indoors and outdoors. These results are
summarized with calendar plots.

## Calendar plots

```{r Outdoor AQHI calendar plot , echo = FALSE, fig.width=5, fig.height=4, fig.align="center"}
generate_AQHI_calendar_plot(outdoor_data_24hr_avg, dataset_col_name = "AQHI",
year_int = year, month_int = month_int, is_outdoor_data = TRUE)

invisible(figure_nums(name = "outdoor_aqhi", caption = sprintf("Calendar plot of outdoor AQHI at your facility during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_aqhi")`

```{r Indoor AQHI calendar plot , echo = FALSE, fig.width=5, fig.height=4, fig.align="center"}
generate_AQHI_calendar_plot(indoor_data_24hr_avg, dataset_col_name = "AQHI",
year_int = year, month_int = month_int, is_outdoor_data = FALSE)

invisible(figure_nums(name = "indoor_aqhi", caption = sprintf("Calendar plot of indoor AQHI at your facility during %s %s.", month_char, year)))
```
`r figure_nums("indoor_aqhi")`

# Nitrogen Dioxide (NO~2~)

## Context

NO~x~ is the combination of nitrogen monoxide (NO) and nitrogen dioxide (NO~2~). Nitrogen monoxide (NO) is formed when nitrogen and oxygen in combustion air combine with one another at the high temperatures experienced in vehicle engines. NO further combines with oxygen to create nitrogen dioxide (NO~2~). In this report we summarize total NO~2~. The effects of NO~2~ exposure include increased inflammation of the airways, coughing and wheezing, reduced lung function, and the onset of asthma and impairment of lung development in children. Metro Vancouver has set air quality recommendations for NO~2~, stating that hourly average exposure to NO~2~ should not exceed 60 ppb, and that annual average exposure to NO~2~ should not exceed 17 ppb. However, because there is no known minimum threshold exposure to (NO~2~) that causes these adverse health impacts, we recommend limiting the exposure to as low as reasonably achievable. 

```{r NOx_setup, include=FALSE}

NO2_diurnals <- get_diurnal_plots(
  outdoor_data, indoor_data,
  "NO2", expression(paste(NO[2], " (ppb)")), NULL
)
```

## Calendar plots

```{r NO2 calendar plot outdoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(outdoor_data, dataset_col_name = "NO2", year_int = year,
  month_int = month_int, title_expression = "Outdoor NO2 Data (ppb)", threshold_value = 42,
  colours = c("green", "green", "yellow", "orange", "red")
)
invisible(figure_nums(name = "outdoor_NO2_calendar", caption = sprintf("Calendar plot of outdoor NO~2~ during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_NO2_calendar")`

```{r NO2 calendar plot indoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(indoor_data, dataset_col_name = "NO2", year_int = year,
  month_int = month_int, title_expression = "Indoor NO2 Data (ppb)", threshold_value = 42,
  colours = c("green", "green", "yellow", "orange", "red")
)
invisible(figure_nums(name = "indoor_NO2_calendar", caption = sprintf("Calendar plot of indoor NO~2~ during %s %s.", month_char, year)))
```
`r figure_nums("indoor_NO2_calendar")`


## Hourly average plots
```{r diurnal_NO2_outdoor, fig.height=3, echo=FALSE}
plot(NO2_diurnals$outdoor_plot)
invisible(figure_nums(name = "outdoor_NO2_diurnal", caption = sprintf("Hourly average plot of outdoor NO~2~ during %s %s.", month_char, year)))
```

`r figure_nums("outdoor_NO2_diurnal")`

```{r diurnal_NO2_indoor, fig.height=3, echo=FALSE}


plot(NO2_diurnals$indoor_plot)
invisible(figure_nums(name = "indoor_NO2_diurnal", caption = sprintf("Hourly average plot of indoor NO~2~ during %s %s.", month_char, year)))
```

`r figure_nums("indoor_NO2_diurnal")`


# Particulate Matter (PM~2.5~)

## Context

PM~2.5~ is the mass fraction of particulate matter with diameters smaller than 2.5 micron. The sources of PM~2.5~ include fuel combustion from traffic and wildfires. Exposure to particulate matter can have effects on childhood health, including irritation of the eyes, nose, and throat, decreased respiratory function, the development of asthma in children. Metro Vancouver's Air Quality Objectives state that 24-hour average exposure to PM~2.5~ should not exceed 25 µg/m^3^, and that annual average exposure to PM~2.5~ should not exceed 8 µg/m^3^. However, we recommend that the exposure to (PM~2.5~) is reduced to as low as reasonably achievable, as there is no known minimum exposure threshold for individuals to experience health impacts.

```{r PM_setup, include=FALSE}

PM_diurnals <- get_diurnal_plots(
  outdoor_data, indoor_data, "PM2_5",
  expression(paste(PM[2.5], " (", µg / m^3, ")")), NULL
)
```

## Calendar plots

```{r PM2.5 calendar plot outdoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(outdoor_data, dataset_col_name = "PM2_5", year_int = year,
  month_int = month_int, title_expression = expression(paste("Outdoor ", PM[2.5], " (", µg / m^3, ")")),
  threshold_value = 25, colours = c("green", "green", "yellow", "red", "#600000")
)

invisible(figure_nums(name = "outdoor_PM_calendar", caption = sprintf("Calendar plot of outdoor PM~2.5~ during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_PM_calendar")`

```{r PM2.5 calendar plot indoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(indoor_data, dataset_col_name = "PM2_5", year_int = year,
  month_int = month_int, title_expression = expression(paste("Indoor ", PM[2.5], " (", µg / m^3, ")")),
  threshold_value = 25, colours = c("green", "green", "yellow", "red", "#600000")
)

invisible(figure_nums(name = "indoor_PM_calendar", caption = sprintf("Calendar plot of indoor PM~2.5~ during %s %s.", month_char, year)))
```
`r figure_nums("indoor_PM_calendar")`


## Hourly average Plots

```{r diurnal_PM_outdoor, fig.height=3, echo=FALSE}
plot(PM_diurnals$outdoor_plot)
invisible(figure_nums(name = "outdoor_PM_diurnal", caption = sprintf("Hourly average plot of outdoor PM~2.5~ during %s %s.", month_char, year)))
```

`r figure_nums("outdoor_PM_diurnal")`

```{r diurnal_PM_indoor, fig.height=3, echo=FALSE}
plot(PM_diurnals$indoor_plot)
invisible(figure_nums(name = "indoor_PM_diurnal", caption = sprintf("Hourly average plot of indoor PM~2.5~ during %s %s.", month_char, year)))
```

`r figure_nums("indoor_PM_diurnal")`


# Carbon Monoxide (CO)

## Context

Carbon monoxide (CO) is a byproduct of incomplete combustion and is associated with the burning of fossil fuels and wildfire smoke events. Exposure to carbon monoxide reduces the body's ability to carry oxygen and can cause headaches, fatigue, dizziness, nausea, and confusion. Metro Vancouver's Air Quality Objectives state that hourly average concentration of CO should not exceed 13 ppm, and that 8-hour average concentration of CO should not exceed 5 ppm.

```{r CO_setup, include=FALSE}
CO_diurnals <- get_diurnal_plots(
  outdoor_data, indoor_data, "CO",
  "CO (ppm)", list(h = 13, lty = 1, lwd = 2, col = "red")
)
```

## Calendar plots

```{r CO calendar plot outdoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(outdoor_data, dataset_col_name = "CO", year_int = year,
  month_int = month_int, title_expression = "Outdoor CO (ppm)", threshold_value = 5,
  colours = c("green", "green", "yellow", "red", "#600000")
)

invisible(figure_nums(name = "outdoor_CO_calendar", caption = sprintf("Calendar plot of outdoor CO during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_CO_calendar")`

```{r CO calendar plot indoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(indoor_data, dataset_col_name = "CO", year_int = year,
  month_int = month_int, title_expression = "Indoor CO (ppm)", threshold_value = 5,
  colours = c("green", "green", "yellow", "red", "#600000")
)

invisible(figure_nums(name = "indoor_CO_calendar", caption = sprintf("Calendar plot of indoor CO during %s %s.", month_char, year)))
```
`r figure_nums("indoor_CO_calendar")`

## Hourly average plots

```{r diurnal_CO_outdoor, fig.height=3, echo=FALSE}
plot(CO_diurnals$outdoor_plot)
invisible(figure_nums(name = "outdoor_CO_diurnal", caption = sprintf("Hourly average plot of outdoor CO during %s %s.", month_char, year)))
```

`r figure_nums("outdoor_CO_diurnal")`

```{r diurnal_CO_indoor, fig.height=3, echo=FALSE}
plot(CO_diurnals$indoor_plot)
invisible(figure_nums(name = "indoor_CO_diurnal", caption = sprintf("Hourly average plot of indoor CO during %s %s.", month_char, year)))
```

`r figure_nums("indoor_CO_diurnal")`


# Ozone (O~3~)

## Context

Ground-level ozone (O~3~) is not a  direct traffic-related emission; however, it is formed from the reaction of nitrogen oxides (NO~x~) and volatile organic compounds (VOCs) (which are traffic-related air pollutants) in the presence of sunlight. Ozone is important in the stratosphere, as it keeps us safe from harmful ultraviolet rays (UV-A and UV-B). However exposure to ground-level ozone can cause shortness of breath or hyperventilation, coughing and/or sore or scratchy throat, inflamed and damaged airways, and aggravation of lung diseases such as asthma, emphysema, and chronic bronchitis. Ground-level ozone is a principal chemical component of smog (smoky fog). Metro Vancouver's Air Quality Objectives state that hourly average exposure to O~3~ should not exceed 82 ppb, and that 8-hour average exposure to O~3~ should not exceed 62 ppb.

In indoor environments, where there is no sunlight, ozone concentrations are generally very low and are therefore not included in this report.

```{r O3_setup, include=FALSE}
O3_diurnals <- get_diurnal_plots(
  outdoor_data, indoor_data, "O3",
  expression(paste(O[3], " (ppb)")),
  list(h = 82, lty = 1, lwd = 2, col = "red")
)
```

## Calendar plot

```{r O3 calendar plot outdoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(outdoor_data, dataset_col_name = "O3", year_int = year,
  month_int = month_int, title_expression = "Outdoor O3 (ppb)", threshold_value = 62,
  colours = c("green", "green", "yellow", "red", "#600000")
)
invisible(figure_nums(name = "outdoor_O3_calendar", caption = sprintf("Calendar plot of outdoor O3 during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_O3_calendar")`

## Hourly average plot

```{r diurnal_O3_outdoors, fig.height=3, echo=FALSE}
plot(O3_diurnals$outdoor_plot)
invisible(figure_nums(name = "outdoor_O3_diurnal", caption = sprintf("Hourly average plot of outdoor O3 during %s %s.", month_char, year)))
```

`r figure_nums("outdoor_O3_diurnal")`


# Carbon dioxoide (CO~2~)

Carbon dioxide (CO~2~) comes from the combustion of organic and inorganic material, as well as respiration from humans. CO~2~ is not typically considered an air pollutant, but it is the primary constituent that contributes to climate change. It can also be a good marker of indoor ventilation. Exposure to high levels of CO~2~ can increase the risk of mucous membrane and respiratory irritation (sore or dry throat, congested nose, sneezing, coughing, and eye irritation), impaired decision-making, and neurophysiological symptoms (headache, tiredness, fatigue, dizziness or difficulty concentrating). Health Canada recommends an individual’s exposure to CO~2~ does not exceed an average of 1000 ppm over a 24-hour period. 

```{r CO2_setup, include=FALSE}
CO2_diurnals <- get_diurnal_plots(
  outdoor_data, indoor_data, "CO2",
  expression(paste(CO[2], " (ppm)")), NULL
)
```

## Calendar plots

```{r CO2 calendar plot outdoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(outdoor_data, dataset_col_name = "CO2", year_int = year,
  month_int = month_int, title_expression = "Outdoor CO2 (ppm)",
  threshold_value = 1000, colours = c("green", "green", "yellow", "red", "#600000")
)
invisible(figure_nums(name = "outdoor_CO2_calendar", caption = sprintf("Calendar plot of outdoor CO2~~ during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_CO2_calendar")`


```{r CO2 calendar plot indoors, echo = FALSE, fig.width=5, fig.height=4, fig.align = "center"}
generate_pollutant_calendar_plot(indoor_data, dataset_col_name = "CO2", year_int = year,
  month_int = month_int, title_expression = "Indoor CO2 (ppm)",
  threshold_value = 1000, colours = c("green", "green", "yellow", "red", "#600000")
)
invisible(figure_nums(name = "indoor_CO2_calendar", caption = sprintf("Calendar plot of indoor CO2~~ during %s %s.", month_char, year)))
```
`r figure_nums("outdoor_CO2_calendar")`

## Hourly average plots

```{r diurnal_CO2_outdoor, fig.height=3, echo=FALSE}
plot(CO2_diurnals$outdoor_plot)
invisible(figure_nums(name = "outdoor_CO2_diurnal", caption = sprintf("Hourly average plot of outdoor CO2~~ during %s %s.", month_char, year)))
```

`r figure_nums("outdoor_CO2_diurnal")`


```{r diurnal_CO2_indoor, fig.height=3, echo=FALSE}
plot(CO2_diurnals$indoor_plot)
invisible(figure_nums(name = "indoor_CO2_diurnal", caption = sprintf("Hourly average plot of indoor CO2~~ during %s %s.", month_char, year)))
```

`r figure_nums("indoor_CO2_diurnal")`